---
title: "Stat333 Project1"
author: "Kelsey Gelden, Brian Johnson, Meishu Zhao, Junyuan Xiang"
date: "10/4/2021"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
We will be attaching our data table "BodyFat" to our code, to analyze and
manipulate it in order to find our best predictors for body fat percentage.
Our goal is to find a simple and robust way to predict our body fat percentage.

```{r}
#BodyFat = read.csv("BodyFat.csv")
#attach(BodyFat)
#summary(BodyFat)
library(readr)
BodyFat <- read_csv("/Users/19kelsog7/Desktop/Fall 2021/STAT333/Stat333_Project1_files/BodyFat.csv")
attach(BodyFat)
summary(BodyFat)
#library(readr)
#BodyFat <- read_csv("Desktop/Fall 2021/STAT333/Stat333_Project1_files/BodyFat.csv")
#View(BodyFat)

```

We noticed, after analyzing our raw data, that some categories are measured in 
different units. We also noticed that one person (#39) takes up the
maximum values for quite a few measurements. We also saw that #42 is said to be
very short (29.5 inches), which seems to be quite inaccurate, especially for 
weighing 205 pounds. We also saw the difference in rounding for each variable.
For example, centimeter measurements are rounded to the nearest tenth, but 
weight and height are rounded to the nearest quarter, and age is rounded to the 
nearest year. These are all things to keep in mind when looking for accuracy. 

```{r}
BodyFat[c(36,39,42,172,182,192,216),]
```

This specific data table of values contain the outliers that we first noticed when
analyzing the original data table. We noticed that our weight distribution, 
along with many circumference measurements will be very skewed to the right due to #39.
There may have also been errors in recording the initial data. Errors could include
simple mistakes and/or inaccurate recordings in density to measure body fat, as it is 
impossible to have negative or 0% body fat. 
We used an estimator to determine how density recorded body fat percentage, and 
the minimum body fat percentage is listed as 0.0%, but the density approximator 
calculated it as -3.6117% body fat. Our maximum body fat percentage could also 
be a result of a miscalculation, as the density approximator determined the body 
fat percentage to be at 47.4874% while the max body fat in the data set is 45.1%, which
is unlikely, but not impossible. We also saw that the height of data point 42 is
extremely low and does not match up with his recorded BMI measurement. We will further
investigate this mis-measurement. 
Our body fat percentage approximator by density is as follows: (495/density)-450 = bf%


```{r}
par(mfrow=c(3,2))
par(mgp=c(1.8,.5,0), mar=c(3,3,1,1))

hist(BODYFAT,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Body Fat %",xlab="Body Fat %")
hist(WEIGHT,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Weight",xlab="Weight (lbs)")
hist(HEIGHT,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Height",xlab="Height (in)")
hist(ABDOMEN,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Abdomen Circumference",xlab="Abdomen Circumference (cm)")
hist(NECK,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Neck Circumference",xlab="Neck Circumference (cm)")
hist(AGE,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Age",xlab="Age (yrs)")
hist(CHEST,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Chest Circumference",xlab="Chest Circumference (cm)")
hist(THIGH,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Thigh Circumference",xlab="Thigh Circumference (cm)")
hist(WRIST,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Wrist Circumference",xlab="Wrist Circumference (cm)")
hist(HIP,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Hip Circumference",xlab="Hip Circumference (cm)")
hist(ANKLE,breaks=100,cex.lab=1.5,cex.main=1.5,
     main="Histogram of Ankle Circumference",xlab="Ankle Circumference (cm)")
```

These histograms explain the distributions of each variable. Body fat percentage 
seems to have a normal distribution, weight is slightly skewed to the right, etc. 
As we can see in these distributions, there are clear outliers in the data set, 
but we are not sure if they are influential points or not. We can study qqplots, 
residual plots, and linear regression models to understand these points' effects 
on this study. Depending on the effect that these points have on our model, we 
will remove the points that seem to have been an inaccurate measurement that 
hinders our prediction model. Also, we realized that male #39 holds the maximum 
value for all measurements except for body fat percentage, height, age, ankle, 
forearm, and density (given that bf% is derived from the density measurement). We 
believe that this is partial proof that there has been a mishap in the data set, 
because #39's measurements are so large, but his body fat percentage is 
significantly less than our maximum of 45.10%, and his weight is 144.15 pounds 
heavier than the person with 45.10% body fat. This data seems to contain 
mis-recordings, but we can check how linear of a relationship there is between 
weight and body fat percentage by plotting a scatter plot. Height also seems to 
be normally distributed except for #42's height of 29.50 inches. For the most part,
circumference measurements seem to be normally distributed, however age is not 
at all. The histogram of the distribution of age shows that there are roughly the 
same amount of people in each age group, but the mid-40's are the most common. 
Let's take a look at some outlier data that we saw from the histograms above.

```{r}
BodyFat[WEIGHT > 300,]
BodyFat[HEIGHT < 60,]
BodyFat[BODYFAT < 4 | BODYFAT > 40,]
BodyFat[ABDOMEN > 130,]
BodyFat[NECK > 45,]
BodyFat[CHEST > 125,]
BodyFat[THIGH > 75,]
BodyFat[HIP > 120,]
BodyFat[ANKLE > 30,]
```
As we analyze these possible outliers, we noticed that, again, #39 is an outlier 
in nearly every measurement. Also, we decided to do our own calculation on #42's 
BMI to test whether his height measurement is accurate or not. According to our 
BMI calculation: [(weight/(height^2))*703 = BMI], with his height and weight, 
his BMI would come to 165.602. This would mean that this person would not be 
living, so we took into account that this height measurement is a mistake, 
therefore we will be removing this row from the data set. We also calculated his 
actual height based off of his BMI and it came out to be about 69.5 inches.
We also decided that a person with 0% body fat would simply not survive, so this 
also must be a miscalculation, and we will be removing data #182. Although 45.1% 
body fat is extreme, we will be keeping this data point as it is still possible 
to live in this condition. In conclusion of these histograms, we will be 
removing data point 182 and 42, and we will take a closer look at data point 39.


```{r}
par(mfrow=c(2,2))
par(mgp=c(1.8,.5,0), mar=c(3,3,1,1))

plot(HEIGHT, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
     xlab = "Height (in)", ylab = "Body Fat %", main = "Body Fat % vs Height")

plot(ANKLE, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Ankle (cm)", ylab = "Body Fat %", main = "Body Fat % vs Ankle Circumference")

plot(AGE, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
     xlab = "Age (yrs)", ylab = "Body Fat %", main = "Body Fat % vs Age")

plot(FOREARM, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Forearm (cm)", ylab = "Body Fat %", main = "Body Fat % vs Forearm Circumference")

plot(WRIST, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Wrist (cm)", ylab = "Body Fat %", main = "Body Fat % vs Wrist Circumference")

plot(BICEPS, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Bicep (cm)", ylab = "Body Fat %", main = "Body Fat % vs Bicep Circumference")

plot(NECK, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
     xlab = "Neck (cm)", ylab = "Body Fat %", main = "Body Fat % vs Neck Circumference")

plot(KNEE, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Knee (cm)", ylab = "Body Fat %", main = "Body Fat % vs Knee Circumference")

plot(THIGH, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Thigh (cm)", ylab = "Body Fat %", main = "Body Fat % vs Thigh Circumference")

plot(WEIGHT, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
     xlab = "Weight (lbs)", ylab = "Body Fat %", main = "Body Fat % vs Weight")

plot(HIP, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Hip (cm)", ylab = "Body Fat %", main = "Body Fat % vs Hip Circumference")

plot(CHEST, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Chest (cm)", ylab = "Body Fat %", main = "Body Fat % vs Chest Circumference")

plot(ABDOMEN, BODYFAT, pch = 19, cex = 0.6, cex.lab = 1.5, cex.main = 1.5,
       xlab = "Abdomen (cm)", ylab = "Body Fat %", main = "Body Fat % vs Abdomen Circumference")
```

While analyzing each of these variables in respect to body fat percentage, we 
noticed that abdomen measurements were most linearly related to body fat 
percentage. We also noticed that height, age, ankle circumference, forearm 
circumference, wrist circumference, and neck circumference were not good 
predictors for body fat percentage, as their relationship between themselves and 
body fat percentage are not significantly linear. We also noticed that biceps 
would not be a reliable predictor of body fat percentage due to its violation of
homoskedasticity. Knee circumference and thigh circumference both start to show 
some linear relationship between themselves and body fat percentage, however we 
will need a more significant relationship between the two to be able to accurately 
predict body fat percentage. Our most significant linear relationships with body 
fat percentage seems to be with chest and abdomen variables according to these 
scatter plots. Therefore, we will further analyze these variables to test for 
body fat percentage. We also concluded that data point 39 is such an extreme 
value, that our prediction model would not be able to properly predict a body 
fat percentage for such an outlier. 
Therefore, we will remove data points 39, 42, and 182.

```{r}
boxplot(ABDOMEN)
```

This box plot of abdomen distribution displays the outlier caused by #39. This is
further proof as to why we believe that we should remove this data point.

```{r}
BodyFat <- BodyFat[-c(42, 182, 39),]
summary(BodyFat)
boxplot(BodyFat$ABDOMEN)
```


This is where we decided to remove our data points that are hindering our
prediction model. We removed data points 39, 42, and 182. Our new summary is 
listed above, along with a box plot of abdomen circumference distribution
without #39 in the data. Outliers still exist in the abdomen distribution, but 
these points (41 and 216) show no inaccurate measurements or hindered recordings, 
therefore we will keep them.

```{r}
## Are there any significant predictors in our set of variables?
BodyFat.lm=lm(BodyFat$BODYFAT~BodyFat$WEIGHT+BodyFat$HEIGHT+BodyFat$HIP+BodyFat$ABDOMEN+BodyFat$AGE+BodyFat$CHEST+BodyFat$NECK+BodyFat$ANKLE+BodyFat$THIGH+BodyFat$BICEPS+BodyFat$WRIST+BodyFat$KNEE+BodyFat$FOREARM+BodyFat$ADIPOSITY,data=BodyFat)
summary(BodyFat.lm)
```

```{r}
chestFit <- lm(BodyFat$BODYFAT ~ BodyFat$CHEST, data = BodyFat)
summary(chestFit)
```

When looking at the chest and body fat percentage summary, we see that there is 
a good amount of significance here. Chest circumference explains about 48% of
variance in body fat percentage. 

```{r}
abdomenFit <- lm(BodyFat$BODYFAT ~ BodyFat$ABDOMEN, data = BodyFat)
summary(abdomenFit)
```

After analyzing this summary, we can clearly see that abdomen circumference is a
good predictor for body fat percentage, as it has a very high F-Statistic, meaning
that it is very significant. Abdomen circumference explains about 67% of variance
in body fat percentage.

```{r}
abdomenChestFit <- lm(BodyFat$BODYFAT ~ BodyFat$ABDOMEN + BodyFat$CHEST, data = BodyFat)
summary(abdomenChestFit)
```

```{r}
reducedModel=lm(BodyFat$BODYFAT~BodyFat$CHEST + BodyFat$ABDOMEN,data=BodyFat)
fullModel= lm(BodyFat$BODYFAT ~ ., data = BodyFat)
anova(reducedModel,fullModel)

reducedModel=lm(BodyFat$BODYFAT~BodyFat$ABDOMEN,data=BodyFat)
fullModel= lm(BodyFat$BODYFAT ~ ., data = BodyFat)
anova(reducedModel,fullModel)
```
This summary explains that chest and abdomen circumference together are significant
predictors for body fat percentage, however the significance is not as high
as the simple predictor of just abdomen measurements, as seen in the smaller F
statistic. The variance of body fat explained by just abdomen circumference alone is 
slightly larger than the variance of body fat percentage explained by chest and 
abdomen circumference together. Our ANOVA tests explain that chest and abdomen together
are significant predictors after accounting for all other variables, but out simple regression model with just abdomen circumference explains the variance very well, and
it is easier to just use one measurement.

In conclusion, we will only use abdomen circumference to predict body fat percentage
in a simple and robust way.


```{r}
##Linearity: Scatter Plot
plot(BodyFat$ABDOMEN,BodyFat$BODYFAT,pch=19,cex=1.2,cex.lab=1.5,cex.main=1.5,
     xlab="Abdomen (cm)",ylab="Body Fat %",main="Body Fat % and Abdomen Circumference")
lmmodel = lm(BODYFAT ~ ABDOMEN, data = BodyFat)
abline(lmmodel,col="blue",lwd=5)
legend("bottomright",c("SLR Line: slope=0.616,intercept=-37.991"),col="blue",lwd=3,lty=c(1))
```

As we take a closer look at this scatter plot of our chosen variable
for linearity, we can say that for every cm increase in abdomen circumference, 
a male would expect to see about a 0.616% increase in body fat. 


```{r}
##Normality: QQ-Plot
qqnorm(rstandard(abdomenFit),pch=19,cex=0.8,cex.lab=1.5,cex.main=1.5,
      main="Normal Q-Q Plot of the Residuals")
abline(a=0,b=1,col="blue",lwd=3)
```

After analyzing our qqplot of abdomen circumference vs body fat percentage,
we can conclude that our abdomen predictor passes the normality assumption. This
is because we see no abnormalities in our plot.

```{r}
##Influential Observations: Cook's Distance

pii = hatvalues(abdomenFit)
cooki = cooks.distance(abdomenFit)

par(mfrow = c(2,1))
n = dim(BodyFat)[1]
plot(1:n,pii,type="p",pch=19,cex=0.6,cex.lab=1.5,cex.main=1.5,
     xlab="Index (Each Observation)",ylab="Pii",main="Leverage Values (Pii)")
plot(1:n,cooki,type="p",pch=19,cex=0.6,cex.lab=1.5,cex.main=1.5,
     xlab="Index (Each Observation)",ylab="Cook's Distance",main="Influence Values (Cook's Distance)")
```

In this diagnostic plot, we can see that data point 216 with bf 45.1% is not 
much of an outlier, because all of his measurements line up with the expected.
Data point 216 is not an influential point. This point's cook's distance is
not significant, therefore we can conclude that it is not an influential
outlier.


```{r}
##Constant error variance, error mean is 0, random error: Residual Plot
res <- resid(abdomenFit)
plot(fitted(abdomenFit), res,pch=19,cex=1,cex.lab=1.5,cex.main=1.5,
    xlab="Predicted Body Fat %", ylab="Standardized Residuals",main="Standardized Residual Plot")
abline(a=0,b=0,col="black",lwd=3)
```

This residual plot displays that the residuals seem scattered evenly and nicely 
around 0 with no significant patterns, after taking out the inaccurate measurements and
influential outliers. Error looks to be randomized around the mean. This allows 
us to pass the assumption of homoscedasticity. 


```{r}
abdomenFit <- lm(BodyFat$BODYFAT ~ BodyFat$ABDOMEN, data = BodyFat)
summary(abdomenFit)
```

With bringing back these statistics, we can build a simple equation to calculate 
our bf%. Our equation is as follows:
0.616(abdomen circumference in cm) - 37.991 = body fat percentage


We can use this in our prediction with an example of 90cm abdomen circumference.
We will obtain our y-value of body fat percentage, and a 95% confidence interval.

```{r}
predict(lmmodel,newdata = data.frame(ABDOMEN=90),interval="predict")
```

This prediction states that if a male has a 90cm abdomen circumference, the 
predicted value of body fat percentage would be 17.44%, and there is a 95% probability
that his body fat percentage would be between about 8.83% and 26.04%. Also, 
according to our prediction model, for every cm increase in abdomen 
circumference, it's expected to increase a male's body fat percentage by 0.616%.

```{r}
predict(lmmodel,newdata=data.frame(ABDOMEN=90),interval="confidence",level=0.95)
```

This confidence interval states that if a male has a 90cm abdomen circumference,
we are 95% confident that average body fat percentages of men who are like him 
are between about 16.880% and 17.998%.

# Summary

Our calculated SLR model to predict Body Fat % comes out to be:

$BodyFatPercentage = -37.991 + (0.616 * Abdomen Circumference(cm))$

**Analysis:** For every 1 centimeter gained in abdomen circumference, a male is expected
to increase in body fat percentage by about 0.616%.

**Example:** For a male within the age range of mid 20's to mid 70's with an abdomen
circumference of 90cm, we predict his body fat percentage to be 17.439%. Along with that,
there would be a 95% probability that this male's body fat percentage falls with in a
range of (8.833, 26.044).

**Relationship:** We find a clear linear relationship between Abdomen circumference and
body fat in males with a significance level of $\alpha = .05$. We attain a p-value of
$2.2e-16$.
This SLR model also explains about 67% of the variations in body fat percentage in males.

**Assumptions:**

1. Linearity: Not violated through scatter plots, random scattering and follows 
the line fairly well.

2. Homoscedasticity: Not violated through residual plots, no fanning or skewness found.

3. Normality: Not violated through QQ plots, hugs 45 degree line, little to no 
skewness of plots.

4. Explains 67% of variability for young adult to adult males, a fairly 
encompassing SLR model.

5. A very low p-value of $2.2e-16$ allows us to draw the conclusion that abdomen 
circumference is useful in predicting Body Fat Percentage.

**Conclusion:** This prediction model shows that with this simple equation, we can
predict body fat percentage of adult males through a single measurement of abdomen
circumference(cm) in a fairly accurate manner. With no assumptions being violated and the
outliers being accounted for, our equation and confidence/prediction intervals for slope
and intercept can give proper estimates, predictions, and ranges from a single abdomen
measurement of circumference.
